BPMN Analyzer 2.0: Instantaneous, Comprehensible, and Fixable Control Flow Analysis for BPMN Models

### 2024-07-05 main.tex
Fixes

We demonstrate instantaneous control flow analysis by benchmarking our tool in \textit{three} different scenarios.
For all our benchmarks, we use the hyperfine benchmarking tool~\cite{peterHyperfine2023} (version 1.18.0), which calculates the average runtime when executing each control flow analysis ten or more times.
The benchmarks were run on Ubuntu 22.04.4 with an AMD Ryzen 7700X processor (4.5GHz) and 32 GB of RAM.
All used BPMN models, our tools to generate them, and benchmarking scripts to run them are available in~\cite{krauterInstantaneousComprehensibleFixable2024}.

\textbf{First}, we benchmarked how our tool handles \textbf{BPMN models of growing size}.
We generated 500 synthetic BPMN models starting with five elements up to 4000 elements, increasing linearly in size.
The models repeatedly contain three activities and an exclusive/parallel block with two branches containing one activity per branch (more details in~\cite{krauterInstantaneousComprehensibleFixable2024}).
The BPMN Analyzer spends from 1 ms up to 9 ms for the BPMN models compared to 0.7 s up to 14 s in our previous tool~\cite{krauterHigherorderTransformationApproach2023}.
In summary, the linear growth of the models leads to a linear growth in state space and, consequently, a linear growth in runtime.

@@

The BPMN Analyzer has comparable BPMN element coverage to~\cite{corradiniFormalApproachAnalysis2021} but is still missing some elements when compared to~\cite{krauterFormalizationAnalysisBPMN2023}, which are currently on the roadmap. 
In addition, we demonstrated the instantaneous analysis of our tool, which outperforms other tools~\cite{krauterFormalizationAnalysisBPMN2023,corradiniFormalApproachAnalysis2021,houhouFirstOrderLogicVerification2022} as described in \autoref{sec:innovations}.
Finally, we have only received positive feedback regarding the tool's usability.

\section{Conclusion \& future work} \label{sec:conclusion}

In this paper, we describe the novel BPMN Analyzer tool that provides instantaneous, comprehensible, and fixable BPMN control flow error detection and is integrated into a popular BPMN modeling tool.
We benchmarked our tool against synthetic and realistic BPMN models to demonstrate instantaneous soundness checking.
We address the three main challenges for providing soundness-checking capabilities to non-expert users as identified in~\cite{fahlandAnalysisDemandInstantaneous2011}.
First, a model checker must be able to check all or most user-created models, i.e., it must support the most used BPMN elements.
This is not a problem for most tools, including ours, which has similar capabilities to~\cite{corradiniFormalApproachAnalysis2021}, and we plan to increase our tool's BPMN coverage as needed.
Second, model checking must be \textit{instantaneous} since long runtimes are unacceptable and often interpreted as tool errors~\cite{fahlandAnalysisDemandInstantaneous2011}.
Third, the biggest challenge for model checking is \textit{consumability}, i.e., reporting the found errors in a comprehensible user interface.
Our tool addresses all these challenges and even provides quick fixes for common errors.
The tool is a BPMN-specific model checker written in Rust paired with an intuitive user interface based on the popular \textit{bpmn.io} ecosystem that allows easy extension in the future.

>>>>

We demonstrate instantaneous control flow analysis by benchmarking our tool in \textit{three} different scenarios.
For all our benchmarks, we use the hyperfine benchmarking tool (version 1.18.0), which calculates the average runtime when executing each control flow analysis ten or more times.
The benchmarks were run on Ubuntu 22.04.4 with an AMD Ryzen 7700X processor (4.5GHz) and 32 GB of RAM (5600 MHz).
All used BPMN models, our tools to generate them, and benchmarking scripts to run them are available in~\cite{krauterInstantaneousComprehensibleFixable2024}.

\textbf{First}, we benchmarked how our tool handles \textbf{BPMN models of growing size}.
We generated 500 synthetic BPMN models starting with five elements up to 4000 elements, increasing linearly in size.
The models repeatedly contain three activities and an exclusive/parallel block with two branches containing one activity per branch (more details in~\cite{krauterInstantaneousComprehensibleFixable2024}).
The BPMN Analyzer spends from 1 ms up to 9 ms for the BPMN models compared to 0.7 s up to 14 s in our previous tool~\cite{krauterFormalizationAnalysisBPMN2023}.
In summary, the linear growth of the models leads to a linear growth in state space and, consequently, a linear growth in runtime.

@@

Our tool has comparable BPMN element coverage to~\cite{corradiniFormalApproachAnalysis2021} but is still missing some elements when compared to~\cite{krauterFormalizationAnalysisBPMN2023}, which are currently on the roadmap. 
In addition, we demonstrated the instantaneous analysis of the BPMN Analyzer, which outperforms other tools~\cite{krauterFormalizationAnalysisBPMN2023,corradiniFormalApproachAnalysis2021,houhouFirstOrderLogicVerification2022} as described in \autoref{sec:innovations}.
Finally, we have only received positive feedback regarding the tool's usability.

\section{Conclusion} \label{sec:conclusion}

In this paper, we describe the novel \textit{BPMN Analyzer} that provides instantaneous, comprehensible, and fixable BPMN control flow error detection and is integrated into a popular BPMN modeling tool.
We benchmarked our tool against synthetic and realistic BPMN models to demonstrate instantaneous soundness checking.
We address the three main challenges: \textit{coverage}, \textit{immediacy}, and \textit{consumability} for providing formal analysis to non-expert users as identified in~\cite{fahlandAnalysisDemandInstantaneous2011}.
In addition, our tool provides quick fixes for common patterns leading to control flow errors.
One can understand the BPMN Analyzer as a BPMN-specific model checker, implemented in Rust paired with an intuitive user interface based on the popular \textit{bpmn.io} ecosystem that allows easy extension in the future.
